---
title: "P8130 HW1"
author: "Chhiring Lama"
date: "2024-09-22"
output: pdf_document
---

```{r setup, include=FALSE, message = FALSE}
library(tidyverse)
library(ggplot2)
library(ggthemes)
```

## Problem 1 (5 points) 

Please classify each of the following variables as qualitative (specify if binary, nominal, or ordinal) or quantitative (specify if discrete or continuous): \

a)	homework feedback, labeled as “poor”, “fair”, “good”, “very good” \
•	Homework feedback labeled as above is a qualitative (ordinal) variable. 


b)	homework feedback, labeled as “fail”, “pass” \
•	Homework feedback here is a qualitative (binary) variable.


c)	country of birth \
•	Country of birth is a qualitative variable. It is a nominal variable as the values here cannot be ordered. 


d)	the quantity of grapes (in lbs) to make 3 liters of wine \
•	Quantity of grapes (in lbs) to make 3 L of wine is a quantitative (continuous) variable. 


e)	number of TAs in the P8130 course \
•	Number of TAs in this course if quantitative (discrete) variable. \

## Problem 2 (15 points) 
In a study of 133 individuals with a recent bike crash history, depression scores were measured using a standardized test. The depression scores for 14 of these individuals are as follows:\

45, 39, 25, 47, 49, 5, 70, 99, 74, 37, 99, 35, 8, 59 \

```{r summary_compute, echo = FALSE}
bike_crash_lst <- c(45, 39, 25, 47, 49, 5, 70, 99, 74, 37, 99, 35, 8, 59)
mean_bike_crash <- mean(bike_crash_lst)
median_bike_crash <- median(bike_crash_lst)
range_bike_crash <- max(bike_crash_lst) - min(bike_crash_lst)
sd_bike_crash <- sd(bike_crash_lst)
```

a)	Compute the following descriptive summaries of these data: mean, median, range, SD.
- Here, number of items (n) = 14, sum of items (S) = $\sum_{i = 1}^{n} x_{i}$ = 45 + 39 + 25 + 47 + 49 + 5 + 70 + 99 + 74 + 37 + 99 + 35 + 8 + 59 = `r sum(bike_crash_lst)`. \

  Mean = S/n = $\sum_{x = 1}^{n} \frac{x_{i}}{n}$ = $\frac{691}{14}$ = `r round(691/14, digits = 2)`. \

  For median, we need to reorder the items in the ascending order so. reordered list = {`r sort(bike_crash_lst)`}. 
  Median = $\frac{(\frac{n}{2})^{th} + (\frac{n+1}{2})^{th}}{2}$ = $\frac{(\frac{14}{2})^{th} + (\frac{14+1}{2})^{th}}{2}$ = $\frac{45 + 47}{2}$.  = $46$.\

  Range = ${max - min}$ = `r max(bike_crash_lst) - min(bike_crash_lst)`.\

  Standard Deviation $(\sigma)$ = $\sqrt{\sigma^2}$ = $\sqrt{\frac{1}{n-1}{\sum_{i = 1}^{n} (x_{i}- \bar{x})^2}}$ where $x$ is an element in the list and $i$ is the index.\
  $\sigma$ = $\sqrt{\frac{1}{14-1}{\sum_{i = 1}^{14} (x_{i}- \frac{691}{14})^2}}$ = `r round(sd_bike_crash, digits = 2)`.\ 

b)	Describe the box plot and the underlying distribution of the data. Use some of the following terms: left-skewed, right-skewed, symmetric, bimodal, unimodal distribution.

```{r plot_dist, message = FALSE, warning = FALSE, fig.width=4, fig.height=4}
bike_crash_list <- list(`Bike Crash` = bike_crash_lst)

boxplot(bike_crash_list)
```

- As shown in the box plot above, the underlying distribution of the data is right-skewed. The distribution is unimodal since the mode is only 1 element (99). However by the nature of the distribution being right-skewed, the median is less the average (mean) and hence the distribution is not symmetric. \

Additionally, 140 individuals with a recent car crash history also participated in the study. The depression scores for 13 of these individuals are given below:\
67, 50, 85, 43, 64, 35, 47, 97, 58, 58, 10, 56, 50 \

a)	Using R, make a side-by-side box plot of the depression scores stratified by type of accident. Make sure you label your figure appropriately.

```{r plot_dist2, message = FALSE, warning = FALSE, fig.width=5, fig.height=4}
car_crash_lst = c(67, 50, 85, 43, 64, 35, 47, 97, 58, 58, 10, 56, 50)

crash_list <- list(`Car Crash` = car_crash_lst,
                  `Bike Crash` = bike_crash_lst)

boxplot(crash_list, main = "Bike and Car Crash History")
```

b)	Describe each of the box plots and the underlying distribution of the data. Use some of the following terms: left-skewed, right-skewed, symmetric, bimodal, unimodal distribution.
- The distribution of bike crash history is right skewed while that of the car crash is symmetric. To check their modes we can use histogram. 

```{r plot_histogram, message = FALSE, warning = FALSE, fig.width=8, fig.height=3}
par(mfrow = c(1, 2))

hist(crash_list$`Car Crash`, main = "Car Crash History", 
     breaks = (2*IQR(crash_list$`Car Crash`)) / 3*sqrt(14),
     col = "lightblue", 
     xlim = c(0, 100))

hist(crash_list$`Bike Crash`, main = "Bike Crash History", 
     breaks = (2*IQR(crash_list$`Bike Crash`)) / 3*sqrt(13), 
     col = "pink", xlim = c(10, 100))

```
  
- The distribution for car crash is a bimodal distribution since there are two modes - 50 and 58 (also shown in the histogram). As mentioned before, distribution for bike crash history is unimodal. \

c) Comparing the 2 box plots, which group appears to have a lower typical depression score?
- Comparing the two groups, group with bike crash history has a lower typical depression score. Both Median (as shown in the box-plot) and mean (55.38 for car crash and 49.36 for bike crash) are lower for bike crash history. 

## Problem 3 (10 points)
Suppose we toss one fair 12-sided die: \

a) Let’s define the event A as “an even number appears”. What is the probability of the event A?
- Here, sample size ($\Omega$) = {1,2, 3, ....., 12} \
  An even number appears ($A$) = {0, 2, 4, 6, 8, 10, 12}\
  In a roll of a dice, the possibilities are all discrete random variable whereby there are total 12 possibility and 6 possibilities for event $A$. Therefore, $P(A)$ = $\frac {6}{12}$ = 0.5. Hence, probability of the event A is 0.5.\

b) Let’s define the event B as “number 10 appears”. What is the probability of the event B?
- $B$ = {10} (only one element) . So, $P(B)$ = 1/12 = `r round(1/12, digits = 2)`. Hence, the probability of the number 10 to appear is `r round(1/12, digits = 2)`. \

c) Compute P(B U A).
- After listing both $A$ and $B$, we know that $B \subset A$. This means $A \cap B$ = $P(B)$. 
  So, \ P(B U A) = $P(A)$ + $P(B)$ - $P(A \cap B)$ = $P(A)$ + $P(B)$ - $P(B)$ = $P(A)$ = 0.5.\
  
d) Are events A and B independent? Why? Prove your answer.
- No, A and B are not independent because if event A and B come from the sample sample space $\Omega$ where $P(A)$ affects $P(B)$ and vice versa.\
  Mathematically, when two events are independent,  $P(A \cap B) = P(A).P(B)$\
  In our case, $P(A).P(B)$ = $(\frac {1}{2}). (\frac {1}{12})$ = $\frac {1}{24}$. However, from c), we know that $P(B \cap A)$ = $P(B)$ = $\frac {1}{12}$. Since $\frac {1}{24} \not= \frac{1}{12}$, A and B are not independent. 

## Problem 4 (10 points)
5% of women above age of 75 have dementia. Among women (75+ years old) with dementia, 80% have positive findings on their CT scan. Among women (75+ years old) who don’t have dementia, 10% will have a positive CT scan findings. A randomly-selected woman (75+ years old) had a positive CT scan findings.\

What is the probability that she actually has dementia? Compute by hand and show the key steps. The answer can be hand written. \

- Here, Probability of dementia in 75+ women ($P(D)$) = 0.05 and probability of healthy cases ($P(H)$) = 1- 0.05 = 0.95 \
        Probability of positive findings in women with dementia = Probability of Positive test given true dementia cases $(P(T^+|D))$ = 0.80 \
        Probability of positive findings in women with dementia = Probability of Positive test given non-dementia cases $(P(T^+|H))$ = 0.80 \
        
  A 75+ years women is already tested positive $(T^+)$, and we want to find out the probability that she has dementia given she tested positive $(P(D|T^+))$. \
  
  Through the requirement of conditional probability, we know that: $(P(D|T^+) = \frac{P(D \cap T^+)}{P(T^+)})$
    
  Using multiplication rule, $P(D \cap T^+)$ = $P(D).P(T^+|D)$, and using total law of probability testing positive is possible either when you have dementia and test positive $(D\cap T^+)$ or when you don't have dementia and test positive $(H\cap T^+)$.\
  
  Replacing this is the main equation:\
  
  $P(D|T^+) = \frac{P(D).P(T^+|D)}{P(D\cap T^+) + P(H\cap T^+)}$ = $\frac{P(D).P(T^+|D)}{P(D). P(T^+|D) + P(H). P(T^+|H)}$ = $\frac{0.05*0.80}{0.05*0.80 + 0.95*0.10}$ = `r round(((0.05*0.80)/((0.05*0.80)+(0.95*0.10))), digits =3)`. \
  
  Hence, the probability that she actually has dementia is `r round(((0.05*0.80)/((0.05*0.80)+(0.95*0.10))), digits =3)`.
  





